Carryover effects in free recall reveal how prior experiences influence memories of new experiences

### 2023-01-11 paper/main.tex
backing up progress on results section: finished describing results from "random" conditions

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage[sort&compress, numbers, super]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{amsopn}

@@

- what do additonal visual features add?  (compare reduced vs. feature rich)
- are the visual features "sticky"?  (compare feature rich vs. reduced (early), also reduced vs. reduced (early))
- are impoverished stimuli "sticky"?  (compare feature rich vs. reduced (late), reduced vs. reduced (late), also reduced (early) vs. reduced (late))

>>>>

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage[sort&compress]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{amsopn}

@@

To obtain a clean preliminary estimate of the consequences on memory of
randomly varying the font colors and locations of presented words (versus
holding the font color fixed at black, and holding the display locations fixed
at the center of the display) we compared participants' performance on the
\textit{feature rich} and \textit{reduced} experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature rich
condition the words' colors and locations varied randomly across words, and in
the reduced condition words were always presented in black, at the center of
the display. Aggregating across all lists for each participant, we found no
difference in recall accuracy for feature rich versus reduced lists ($t(126) =
-0.290, p = 0.772$). However, participants in the feature rich condition
clustered their recalls substantially more along every dimension we examined
(temporal clustering: $t(126) = 10.624, p < 0.001$; category clustering:
$t(126) = 10.077, p < 0.001$; size clustering: $t(126) = 11.829, p < 0.001$;
word length clustering: $t(126) = 10.639, p < 0.001$; first letter clustering:
$t(126) = 7.775, p = 0.000$; see \textit{Permutation-corrected feature
clustering scores} for more information about how we quantified each
participant's clustering tendencies.) Taken together, these comparisons suggest
that adding new features changes how participants organize their memories of
studied words, even when those new features are independent of the words
themselves and even when the new features vary randomly across words. We found
no evidence that those additional uninformative features were distracting (in
terms of their impact on memory performance), but they did affect participants'
recall dynamics (measured via their clustering scores).

We also wondered whether adding these irrelevant visual features to later lists
(after the participants had already studied impoverished lists), or removing
the visual features from later lists (after the participants had already
studied visually diverse lists) might affect memory performance. In other
words, we sought to test for potential effects of changing the ``richness'' of
participants' experiences over time. All participants studied and recalled a
total of 16 lists; we defined \textit{early} lists as the first eight lists and
\textit{late} lists as the last eight lists each participant encountered. To
help interpret our results, we compared participants' memories on early versus
late lists in the above feature rich and reduced conditions. Participants in
both conditions remembered more words on early versus late lists (feature rich:
$t(66) = 4.553, p < 0.001$; reduced: $t(60) = 2.434, p = 0.018$). Participants
in the feature rich (but not reduced) conditions exhibited more temporal
clustering on early versus late lists (feature rich: $t(66) = 2.318, p =
0.024$; reduced: $t(60) = 0.929, p = 0.357$). And participants in both
conditions exhibited more semantic (category and size) clustering on early
versus late lists (feature rich, category: $t(66) = 3.805, p < 0.001$; feature
rich, size: $t(66) = 2.190, p = 0.032$; reduced, category: $t(60) = 2.856, p =
0.006$; reduced, size: $t(60) = 2.947, p = 0.005$). Participants in the reduced
(but not feature rich) conditions exhibited more lexicographic clustering on
early versus late lists (feature rich, word length: $t(66) = 0.161, p = 0.872$;
feature rich, first letter: $t(66) = 0.410, p = 0.683$; reduced, word length:
$t(60) = 3.528, p = 0.001$; reduced, first letter: $t(60) = 2.275, p = 0.026$).
Taken together, these comparisons suggest that even when the presence or
absence of irrelevant visual features is stable across lists, participants
still exhibit some differences in their performance and memory organization
tendencies for early versus late lists.

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the irrelevant visual features constant on early
lists, but allowed them to vary randomly on late lists. In a \textit{reduced
(late)} condition, we allowed the irrelevant visual features to vary randomly
on early lists, but held them constant on late lists. Given our above findings
that (a) participants tended to remember more words and exhibit stronger
clustering effects on feature rich (versus reduced) lists, and (b) participants
tended to remember more words and exhibit stronger clustering effects on early
(versus late) lists, we expected these differences to be enhance in the reduced
(early) condition and diminished in the reduced (late) condition. However, to
our surprise, participants in \textit{neither} condition exhibited reliable
early versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
= 0.278$), nor feature based clustering (reduced (early), category: $t(41) =
0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
0.966, p = 0.340$). We hypothesized that adding or removing the irrelevant
features was acting as a sort of ``event boundary'' between early and late
lists. In prior work, we (and others) have found that memories formed just
after event boundaries can be enhanced~\citep[e.g.,][]{MannEtal16}.

% JRM: STOPPED HERE...
